---
title: "Epidemiological Informatics: Modeling STI Screening Utilization Gaps"
subtitle: "KNUST Public Health Informatics Working Briefing"
author: "Lead Portfolio Engineer: Afriyie Albert Dwamena (Physician Assistant)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme: bootstrap
    highlight: tango
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    highlight: tango
---

```{r setup, include=FALSE}
# 1. Establish deterministic environment configuration
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center')

# 2. Inject core public health and data visualization packages
library(tidyverse)
library(epitools)
library(knitr)
```

\newpage

# Executive Research Abstract

This working briefing compiles cross-sectional epidemiological insights derived from a cohort of
undergraduate students ($N = 333$) at the Kwame Nkrumah University of Science and Technology
(KNUST). By combining relational database extraction with supervised machine learning and
biostatistical inference, this pipeline isolates the institutional awareness disparities and
behavioural friction vectors driving Sexually Transmitted Infection (STI) screening utilisation
gaps.

---

# Downstream Data Ingestion

We load the programmatic data array exported directly from our automated Python pipeline layer:

```{r load_data}
# Read the processed cohort dataset
cohort_data <- read.csv("../data/processed_cohort_data.csv")

# Display structural baseline verification
dim(cohort_data)
```

The underlying dataset contains exactly `r nrow(cohort_data)` records and
`r ncol(cohort_data)` feature columns, spanning demographic variables, academic identifiers,
clinical risk-perception matrices, and historic health-seeking behaviours.

---

# Epidemiological Risk Analytics & Odds Ratios (OR)

To evaluate the impact of biological sex on active clinical diagnostic utilisation, we construct
an epidemiological contingency matrix and compute Wald Odds Ratios with exact 95% confidence
intervals.

```{r odds_ratio_computation}
# 1. Construct the 2x2 cross-tabulation matrix for biological sex vs screening history
contingency_table <- table(cohort_data$gender, cohort_data$ever_tested)

# 2. Run the Wald Odds Ratio test via the epitools framework
wald_results <- oddsratio.wald(contingency_table)

# 3. Extract and display exact metrics
print(wald_results$measure)
```

---

# Machine Learning Model Diagnostic Ingestion

Rather than executing loose, unverified figures, we dynamically pull the exact diagnostic
evaluation charts generated by our `scikit-learn` regularised Logistic Regression classifier
script (`notebooks/02_predictive_modeling.ipynb`):

```{r model_diagnostics, echo=FALSE, out.width="85%", fig.cap="ROC Curve — Regularised Logistic Regression Classifier"}
# Dynamically pull the model diagnostic plot into the document compiler
roc_paths <- c("../notebooks/model_roc_curve.png", "model_roc_curve.png")
found_roc  <- roc_paths[file.exists(roc_paths)]

if (length(found_roc) > 0) {
  knitr::include_graphics(found_roc[1])
} else {
  cat("⚠️  Diagnostic visualisation [model_roc_curve.png] pending execution in Notebook 2.")
}
```

---

# Summary of Relational EHR Database Extractions

The 3rd Normal Form (3NF) relational SQL database instance (`database/knust_health.db`)
computes the following institutional metrics across academic sectors:

- **The Intention–Behaviour Action Gap:** Marked drop-offs in diagnostic screening utilisation
  are isolated within the Engineering and Science faculties despite baseline clinic awareness.
  This suggests that accessibility barriers and scheduling frictions override positive intent.

- **Cognitive Risk Insulation:** Advanced window-function partitioning establishes that
  partnered student demographics within the College of Science maintain the lowest overall
  peer-risk perception, indicating a relationship-driven cognitive shield that delays active
  screening.

---

*Document Status: Compiled and Verified*  
*Informatics Framework: R Markdown + Pandoc Pipeline*
